Learning from Massive Noisy Labeled Data for Image Classification

### 2015-04-11 paper.tex
Polish punctuations

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}

@@

%%%%%%%%% TITLE
\title{Learning From Massive Noisy Labeled Data for Image Classification}
\author[1]{Tong Xiao}
\author[2]{Tian Xia}
\author[2]{Yi Yang}
\author[2]{Chang Huang}
\author[1]{Xiaogang Wang}
\affil[1]{The Chinese University of Hong Kong}
\affil[2]{Baidu Research}
\renewcommand\Authands{, and }

@@

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. We manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

@@

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Figure~\ref{fig:compare_noise_level}.

@@

To illustrate the relations between noisy and true labels, we derive their conditional probability from Eq~\ref{eq:joint_distribution}
\begin{equation}
    p(\tilde{\vec{y}} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}}, \vec{z} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{z} | \vec{x})
\end{equation}
which can be interpreted as a mixture model. Given an input image $\vec{x}$, the conditional probability $p(\vec{z} | \vec{x})$ can be seen as the prior of each mixture component. This makes a key difference between our work and~\cite{sukhbaatar2014learning}, where they assume $\tilde{\vec{y}}$ is conditionally independent with $\vec{x}$ if $\vec{y}$ is given. All the images share a same noise model in~\cite{sukhbaatar2014learning}, while in our approach each data sample has its own.

@@

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood,
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using the current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

@@

\subsection{Estimating Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts in Eq~\eqref{eq:joint_distribution} could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid such degeneration, we estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we discard those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

Next, we will show that the problem can be solved by a simple yet efficient algorithm in $O(N'+L^2)$ time complexity. Denote the optimal solution by $\mat{C}^*$ and $\vec{t}^*$
\begin{theorem}
$\mat{C}^*_{ij} \neq 0 \Rightarrow \mat{C}^*_{ij} > \alpha, \forall i,j \in \{1,\dots,L\}$
\end{theorem}
\begin{proof}
Suppose there exists some $i,j$ such that $0 < \mat{C}^*_{ij} \leq \alpha$. Then we conduct the following operations. First, we set $\mat{C}^*_{ij}=0$ while adding its original value to other elements in column $j$. Second, for all the samples $n$ where $\tilde{\vec{y}}^{(n)}_i=1$ and $\vec{y}^{(n)}_j=1$, we set $\vec{t}_n$ to 1. The resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$(\tilde{\vec{y}}^{(i)},\vec{y}^{(i)})=(\tilde{\vec{y}}^{(j)},\vec{y}^{(j)}) \Rightarrow \vec{t}^*_i=\vec{t}^*_j, \forall i,j \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
Suppose $\tilde{\vec{y}}^{(i)}_k=\tilde{\vec{y}}^{(j)}_k=1$ and $\vec{y}^{(i)}_l=\vec{y}^{(j)}_l=1$ but $\vec{t}^*_i \neq \vec{t}^*_j$. From Theorem 1 we know that elements in $\mat{C}^*$ is either 0 or greater than $\alpha$. If $\mat{C}^*_{kl}=0$, we can set $\vec{t}^*_i=\vec{t}^*_j=1$, otherwise we can set $\vec{t}^*_i=\vec{t}^*_j=0$. In either case the resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} > \alpha \Leftrightarrow \vec{t}^*_i=0$ and $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0 \Leftrightarrow \vec{t}^*_i=1, \forall i \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
The first part is straightforward. For the second part, $\vec{t}^*_i=1$ implies $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} \leq \alpha$. By using Theorem 1 we know that $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0$.
\end{proof}

@@

Due to the rearrangement inequality, we can prove that in the optimal solution,
\begin{equation}
    \max(\alpha,\vec{c}^*_1) \geq \max(\alpha,\vec{c}^*_2) \geq \cdots \geq \max(\alpha,\vec{c}^*_L).
\end{equation}
Then by using Theorem 3, there must exist a $k^* \in \{1,\dots,L\}$ such that
\begin{equation} \label{eq:opt_t}
\begin{aligned}
    &\vec{t}^*_i = 0, i = 1,\dots,k^* \\
    &\vec{t}^*_i = 1, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}
This also implies that only the first $k^*$ elements of $\vec{c}^*$ have nonzero values (greater than $\alpha$ actually). Furthermore, if $k^*$ is known, finding the optimal $\vec{c}^*$ is to solve the following problem:
\begin{equation}
\begin{aligned}
\max_{\vec{c}} \quad & E=\sum_{k=1}^{k^*} m_k \log \mat{c}_k,\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{k^*} \vec{c}_k = 1,
\end{aligned}
\end{equation}
whose solution is
\begin{equation} \label{eq:opt_c}
\begin{aligned}
    &\vec{c}^*_i = \frac{m_i}{\sum_{k=1}^{k^*}{m_k}}, i = 1,\dots,k^*,\\
    &\vec{c}^*_i = 0, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}

The above analysis leads to a simple algorithm. We enumerate $k^*$ from 1 to $L$. For each $k^*$, $\vec{t}^*$ and $\vec{c}^*$ are computed by using Eq~\eqref{eq:opt_t} and~\eqref{eq:opt_c}, respectively. Then we evaluate the objective function $E$ and record the best solution.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Figure~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq~\eqref{eq:gradient}.

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[nocompress]{cite}

@@

%%%%%%%%% TITLE
\title{Learning from Massive Noisy Labeled Data for Image Classification}
\author[1]{Tong Xiao}
\author[2]{Tian Xia}
\author[2]{Yi Yang}
\author[2]{Chang Huang}
\author[1]{Xiaogang Wang}
\affil[1]{The Chinese University of Hong Kong}
\affil[2]{Baidu Research}
\renewcommand\Authands{, and }

@@

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Fig.~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. We manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Fig.~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Fig.~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Fig.~\ref{fig:overview}).
\end{itemize}

@@

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Fig.~\ref{fig:noise_pgm}.

@@

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$.
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$.
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$.
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1,
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Fig.~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$,
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x}).
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Fig.~\ref{fig:compare_noise_level}.

@@

To illustrate the relations between noisy and true labels, we derive their conditional probability from Eq.~\ref{eq:joint_distribution},
\begin{equation}
    p(\tilde{\vec{y}} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}}, \vec{z} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{z} | \vec{x}),
\end{equation}
which can be interpreted as a mixture model. Given an input image $\vec{x}$, the conditional probability $p(\vec{z} | \vec{x})$ can be seen as the prior of each mixture component. This makes a key difference between our work and~\cite{sukhbaatar2014learning}, where they assume $\tilde{\vec{y}}$ is conditionally independent with $\vec{x}$ if $\vec{y}$ is given. All the images share a same noise model in~\cite{sukhbaatar2014learning}, while in our approach each data sample has its own.

@@

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood,
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}.
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using the current parameters $\theta^{(t)}$,
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}.
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta).
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2).
\end{multline}

@@

\subsection{Estimating Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq.~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts in Eq.~\eqref{eq:joint_distribution} could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid such degeneration, we estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)}).
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we discard those samples and reinterpret the problem in another form by exploiting Eq.~\eqref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i},\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L, \\
                    & \vec{t} \in \{0,1\}^{N'},
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

Next, we will show that the problem can be solved by a simple yet efficient algorithm in $O(N'+L^2)$ time complexity. Denote the optimal solution by $\mat{C}^*$ and $\vec{t}^*$
\begin{theorem}
$\mat{C}^*_{ij} \neq 0 \Rightarrow \mat{C}^*_{ij} > \alpha, \forall i,j \in \{1,\dots,L\}$.
\end{theorem}
\begin{proof}
Suppose there exists some $i,j$ such that $0 < \mat{C}^*_{ij} \leq \alpha$. Then we conduct the following operations. First, we set $\mat{C}^*_{ij}=0$ while adding its original value to other elements in column $j$. Second, for all the samples $n$ where $\tilde{\vec{y}}^{(n)}_i=1$ and $\vec{y}^{(n)}_j=1$, we set $\vec{t}_n$ to 1. The resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$(\tilde{\vec{y}}^{(i)},\vec{y}^{(i)})=(\tilde{\vec{y}}^{(j)},\vec{y}^{(j)}) \Rightarrow \vec{t}^*_i=\vec{t}^*_j, \forall i,j \in \{1,\dots,N'\}$.
\end{theorem}
\begin{proof}
Suppose $\tilde{\vec{y}}^{(i)}_k=\tilde{\vec{y}}^{(j)}_k=1$ and $\vec{y}^{(i)}_l=\vec{y}^{(j)}_l=1$ but $\vec{t}^*_i \neq \vec{t}^*_j$. From Theorem 1 we know that elements in $\mat{C}^*$ is either 0 or greater than $\alpha$. If $\mat{C}^*_{kl}=0$, we can set $\vec{t}^*_i=\vec{t}^*_j=1$, otherwise we can set $\vec{t}^*_i=\vec{t}^*_j=0$. In either case the resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} > \alpha \Leftrightarrow \vec{t}^*_i=0$ and $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0 \Leftrightarrow \vec{t}^*_i=1, \forall i \in \{1,\dots,N'\}$.
\end{theorem}
\begin{proof}
The first part is straightforward. For the second part, $\vec{t}^*_i=1$ implies $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} \leq \alpha$. By using Theorem 1 we know that $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0$.
\end{proof}

@@

Due to the rearrangement inequality, we can prove that in the optimal solution,
\begin{equation}
    \max(\alpha,\vec{c}^*_1) \geq \max(\alpha,\vec{c}^*_2) \geq \cdots \geq \max(\alpha,\vec{c}^*_L).
\end{equation}
Then by using Theorem 3, there must exist a $k^* \in \{1,\dots,L\}$ such that
\begin{equation} \label{eq:opt_t}
\begin{aligned}
    &\vec{t}^*_i = 0, i = 1,\dots,k^*, \\
    &\vec{t}^*_i = 1, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}
This also implies that only the first $k^*$ elements of $\vec{c}^*$ have nonzero values (greater than $\alpha$ actually). Furthermore, if $k^*$ is known, finding the optimal $\vec{c}^*$ is to solve the following problem:
\begin{equation}
\begin{aligned}
\max_{\vec{c}} \quad & E=\sum_{k=1}^{k^*} m_k \log \mat{c}_k,\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{k^*} \vec{c}_k = 1,
\end{aligned}
\end{equation}
whose solution is
\begin{equation} \label{eq:opt_c}
\begin{aligned}
    &\vec{c}^*_i = \frac{m_i}{\sum_{k=1}^{k^*}{m_k}}, i = 1,\dots,k^*,\\
    &\vec{c}^*_i = 0, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}

The above analysis leads to a simple algorithm. We enumerate $k^*$ from 1 to $L$. For each $k^*$, $\vec{t}^*$ and $\vec{c}^*$ are computed by using Eq.~\eqref{eq:opt_t} and~\eqref{eq:opt_c}, respectively. Then we evaluate the objective function $E$ and record the best solution.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Fig.~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq.~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq.~\eqref{eq:gradient}.

Directly training the whole network with random initialization is impractical, because the posterior computation could be totally wrong. Therefore, we need to pretrain each CNN component with strongly supervised data. Images and their ground truth labels in the dataset $\mathcal{D}_c$ are used to train the CNN that predicts $p(\vec{y}|\vec{x})$. On the other hand, the optimal solutions of $\vec{z}^{(1)},\cdots,\vec{z}^{(N)}$ in Eq.~\eqref{eq:opt_all} are used to train the CNN that predicts $p(\vec{z}|\vec{x})$.
